\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\begin{document}
\subsection{MFE, retake}

Part A.

\begin{enumerate}
\item Give an example of 
\begin{enumerate}
\item a function $f(x,y)$ that has a non-zero gradient in all the points except the point $(0,5)$ and zero gradient in the point $(0,5)$
\item a function $g(x,y)$ that has no gradient on the line $x=3y$ and a non-zero gradient when $x\neq 3y$
\end{enumerate}

\item The population of a certain country grows exponentially, $N_t= N_{1990}\cdot \exp (r(t-1990))$. The population was 70 million in 1990 and 80 million in 2000,
what will be the population in 2014?

\item Find and classify the extrema of the function $f(x,y)=x^2-y^2$ subject to $x^2+y^2=1$.

\item Given the system
\[
\begin{cases}
xe^{u+v}+2uv=1 \\
ye^{u-v}-\frac{u}{1+v}=2x
\end{cases}
\]
find $du$ and $dv$ at $x_0 = 1$, $y_0 = 2$, $u_0 = 0$, $v_0 = 0$.



\item Consider the objective function $f(x,y)=4kx^3+k^2xy+3ky^4-13x-13y$. The point $(x,y)=(1,1)$ is the maximum of the function. Find the value of $k$

\item In the macroeconomic linear IS-LM model for the closed economy
$Y=\bar{C}+m(Y-T)+G+\bar{I}-ar$ and $\bar{L}+bY-cr=M_s$, where $M_s$ is money supply, $r$ --- interest rate, $G$ --- government expenditures, $T$ --- lump sum tax and the constant parameters $\bar{C}>0$, $0<m<1$, $\bar{I}>0$, $a>0$, $\bar{L}>0$, $b>0$, $c>0$. Find the formulas for $dr/dT$, $dY/dT$. Assume that government expenditures and money supply are fixed exogenous variables.
\end{enumerate}


Part B.

\begin{enumerate}[resume]
\item The production function of a firm is given by $y=\sqrt{x_1}+\sqrt{x_2}$, where $x_1$ and $x_2$ are the factors of production. Given the factor prices $w_1=5w$, $w_2=w>0$ find the total costs function of the firm.
\item A consumer splits her time $\bar{L}$ hours a week between labor and leisure. Her utility function is represented by $u(c,l)=c^{\alpha}l^{1-\alpha}$, where $c$ is the amount of consumption and $l$ is leisure in hours and $0<\alpha<1$. The weekly budget constraint is written as $pc+wl=w\bar{L}$,
where $p$ is the price of consumption, $w$ is the hourly wage rate.
\begin{enumerate}
\item Find the consumer’s optimal bundle $(c^*,l^*)$. Justify your answer by checking second-order conditions or otherwise.
\item Let $\alpha=1/4$, $\bar{L}=168$, $p=10$, $w=5$. Using Envelope Theorem estimate the change in the maximum value of her utility if the wage rate has decreased by $0.5$.
\end{enumerate}

\end{enumerate}

\end{document}