\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=1.5cm,right=2.5cm,top=1.5cm,bottom=1.5cm]{geometry}
\begin{document}
\subsection*{Assignment 23 (20 April)}

\begin{enumerate}



\item Solve the following differential-equation system and analyse the time path:
\[
\begin{cases}
x'(t)-x(t)-12y(t)=-60 \\
y'(t)+x(t)+6y(t)=36
\end{cases}
\]
with $x(0)=13$, $y(0)=4$.

\item  Solve the following difference-equation system and analyse the time path
\[
\begin{cases}
x_{t+1}+x_t+2y_t=24 \\
y_{t+1}+2x_t-2y_t=9
\end{cases}
\]
 with $x_0=10$, $y_0=9$.

\item  Find the eigenvalues of the matrix 
\[
A=\begin{pmatrix}
4 & -1 & -1 \\ 
1 & 2 & -1 \\ 
1 & -1 & 2
\end{pmatrix} 
\]
and find the eigenvectors corresponding to each eigenvalue. Hence find an invertible matrix $P$  and a diagonal matrix $D$ such that $D=P^{-1}AP$. 

\begin{enumerate}
\item  Use your result to find the sequences $x_t$, $y_t$, $z_t$, such that $x_0=1$, $y_0=3$, $z_0=2$ and for $t>0$,
\[
\begin{cases}
x_{t+1}=4x_t-y_t-z_t \\
y_{t+1}=x_t+2y_t-z_t \\
z_{t+1}=x_t-y_t+2z_t
\end{cases}
\] 


\item Use your result to find the functions $x(t)$, $y(t)$, $z(t)$  such that  
$x(0)=1$, $y(0)=3$, $z(0)=2$ and 
\[
\begin{cases}
x'=4x-y-z \\
y'=x+2y-z \\
z'=x-y+2z
\end{cases}
\] 



\end{enumerate}
\item  Given the pay-off bi-matrix 


\begin{tabular}{c|cccc}
 & e & f & g & h \\ 
\hline
a & 1,1 & 5,3 & 3,2 & 3,4 \\ 
b & 1,0 & 1,2 & 2,5 & 2,6 \\ 
c & 3,3 & 3,5 & 3,4 & 4,4 \\ 
d & 2,2 & 0,3 & 2,2 & 2,2 \\ 
\end{tabular} 

Find all Nash equilibria in pure and mixed strategies. Check whether they are Pareto-optimal.

\item Given the pay-off matrix of a zero-sum game

\begin{tabular}{c|ccc}
 & c & d & e \\ 
\hline 
a & -1 & -3 & 0 \\ 
b & -4 & 2 & -5 \\ 
\end{tabular} 

Find all Nash equilibria in pure and mixed strategies. Check whether they are Pareto-optimal.

\end{enumerate}

\newpage
\subsection*{Assignment 24 (27 April)}

\begin{enumerate}

\item A policymaker desires to double in 10 periods of time the value of GDP $Y_t$ produced in period $t$. Evolution of GDP over time is given by equation $2Y_{t+2}-3Y_{t+1}+Y_t=2^t+t$. Is doubling of GDP feasible? If the answer is positive,  find the period $t$ when the value of $Y_t$ will first exceed $2Y_0$, where $Y_0$ is the initial GDP.

\item  Solve the system of ODE
\[
\begin{cases}
\dot{x}=3x-2y \\
\dot{y}=2x-y
\end{cases}
\]




\item  Let production function $F(K, L)$ be twice continuous differentiable and homogeneous of the first degree. Show that its Hessian matrix has a zero determinant.

\item  Two candidates, A and B, compete in an election. Of the 100 citizens, $k$ support candidate A and $m= 100 - k$ support candidate B. Each citizen decides whether to vote, at a cost, for the candidate she supports, or to abstain. A citizen who abstains receives the payoff of 2 if the candidate she supports wins, 1 if this candidate ties for first place, and 0 if this candidate loses. A citizen who votes receives the payoffs $2 - c$, $1 - c$, and $-c$ in these three cases, where $0 < c < 1$.
\begin{enumerate}
\item  For $k = 50$, find the set of Nash equilibria in pure strategies. (Is the action profile in which everyone votes a Nash equilibrium? Is there any Nash equilibrium in which the candidates tie and not everyone votes? Is there any Nash equilibrium in which one of the candidates wins by one vote? Is there any Nash equilibrium in which one of the candidates wins by two or more votes?)
\item  What is the set of Nash equilibria in pure strategies for $k < 50$?
\end{enumerate} 

\item General A is defending territory accessible by two mountain passes against an attack by general B. General A has three divisions at her disposal, and general B has two divisions. Each general allocates her divisions between the two passes. General A wins the battle at a pass if and only if she assigns at least as many divisions to the pass as does general B; she successfully defends her territory if and only if she wins the battle at both passes. Find all the mixed strategy equilibria.


\end{enumerate}

\newpage
\subsection*{Additional Computer Home Assignment (27 April)}

You may use any open source software, R is recommended but not mandatory. Please provide not only the answers but also the code.  

\begin{enumerate}

\item Find all the eigenvalues and eigenvectors of the following matrix

\[
\begin{pmatrix}
5 & 2 & -1\\
2 & 3 & 6\\
-1 & 6 & -2
\end{pmatrix}
\]

\item Draw the solution of the second order equation

\[
y''+(x+1)\cdot y' + y=\arctan x
\]

with initial conditions $y(0)=0$, $y'(0)=1$.

\item Find numerically the global minimum of the function $f(x,y)=x^4+y^8+2xy-4x+xy^2$.

\item Find at least one Nash Equilibrium of the following zero-sum game:

\begin{tabular}{c|ccc}
 & e & f & g \\ 
\hline 
a & 2 & -3 & 0 \\ 
b & 0 & 3 & -5 \\ 
c & -4 & 0 & 2 \\ 
\end{tabular} 

\item Masha and Sasha play the following game. Masha writes two numbers on two small sheets of paper. The number on the first sheet is uniformly distributed on $[0;1]$, the number on the second sheet is just the first number squared. Sasha selects one of the two sheets at random with equal probabilities. He looks at the number and should guess whether he has selected the biggest number or not. Find his optimal strategy. What is the probability that Sasha will win if he uses the optimal strategy?

\end{enumerate}




\end{document}