\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}
\section*{MOR. Retake --- 15.09.2015}

\pagestyle{empty}

\begin{enumerate}
\item A problem with the mixed constraints is given: $3x_1 x_2-x_2^3 \to \max$

subject to $2x_1+5x_2 \geq 20$, $x_1-2x_2=5$, $x_1, x_2 \geq 0$.
\begin{enumerate}
\item Check NDCQ conditions.
\item  Form the Langrangian function.
\item Solve the maximization problem by the use of Langrange method or otherwise.
\item Justify that the maximum point was found.
\end{enumerate}

\item  Consider the linear programming problem with parameter $\beta$ and nonnegative $x_i$:
\begin{align*}
3x_1+x_2-x_3+4x_4 \to \max \\
x_1+x_2-x_3+x_4\leq 1 \\
\beta x_1+x_2+x_3+2x_4 \leq 2
\end{align*}

\begin{enumerate}
\item Find the optimal values of the primal variables for $\beta=6$.
\item Find the function $\phi(\beta)$, where $\phi(\beta)$  is the maximum value of the objective function for fixed value of $\beta$.
\item Sketch the graph of $\phi(\beta)$.
\end{enumerate}
\item Find the general solution of the differential equation $y''+6y'+9y=xe^{-2x}+\sin(x)$
\item Consider the system of difference equations
$$ 
\left\{ \begin{array}{l}
x_{t+1}=2x_t-4y_t \\
y_{t+1}=x_t-3y_t+3
\end{array} \right.
$$
\begin{enumerate}
\item Solve the system
\item Find the equilibrium solution and check whether it's stable
\end{enumerate}


\item Find all pure and mixed Nash equilibria in the following bimatrix game:


\begin{tabular}{c|ccc}
 & d & e & f \\ 
\hline 
a & 4;5 & 1;4 & 1;1  \\ 
b & 2;8 & 5;0 & 0;4  \\ 
c & 0;3 & 2;2 & 5;7  \\ 
\end{tabular} 


\item There is an auction of a painting with two players. The value of the painting for the first player is a random variable $v_1$, for the second player --- $v_2$. The random variables $v_1$ and $v_2$ are independent and identically distributed from 0 to 1 million dollars with density function $f(t)=2t$. Each player makes the bid $b_i$ knowing only his own value of the painting. The player who makes the highest bid gets the painting and pays his bid. 

Find a Nash equilibrium where each player uses linear strategy of the form $b_i=k\cdot v_i$.


\end{enumerate}



\end{document}