\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage[left=1.5cm, right=1.5cm, top=1.5cm, bottom=2cm]{geometry}

\def \RR{\mathbb{R}}

\begin{document}
\thispagestyle{empty}
\textbf{Variant 1}. 2017-09-21. Please, don't forget to write you variant number. Sections A and B will make up 60\% and 40\% of the exam grade, respectively. Total duration of the exam is 120 min. Good luck! :)

% две оптимизации - кроме лин
% линейное программировани
% теория игр бруталка

\textbf{SECTION A}

\begin{enumerate}

\item Find indefinite integrals
\begin{enumerate}
  \item  $\int {{e^{2x}}\cos^2 x \, dx}$;
  \item  $\int {\frac{{x - 3}}{{{x^2} - 6x + 9}} \, dx}$.
\end{enumerate}

\item Solve the differential equation
\[
y''' - y'' + 6y' - 6 = 42\sin (x\sqrt 6 ).
\]

\item Solve the difference equation
\[
{y_{t + 2}} - 8{y_{t + 1}} + 16{y_t} = 5t.
\]

\item The function $f(x, y)$ is non-constant and homogeneous. It is also known that $h(x, y) = f'_x(x, y) + 3x^2y$ is homogeneous of degree 3. Find the value of $\frac{xf'_x(x, y) + yf'_y(x, y)}{f(x, y)}$.

\item Solve the following linear programming problem:
\[
\begin{cases}
2x_1 + 2x_2 + 3x_3 \to \min \\
x_1 \geq 0, x_2 \geq 0, x_3 \geq 0 \\
3x_1 + 5x_2 + x_3 \geq 8 \\
5x_1 + 3x_2 + x_3 \geq 9 \\
\end{cases}.
\]

\item Maximize the function
\[
11 + 10x_1 - x_1^2 -3x_2 + 8x_3 - x_3^2
\]
subject to constraints $2x_1 -x_2+4x_3 \leq 10$ and $x_2 \leq 100$.

\end{enumerate}

\textbf{SECTION B}

\begin{enumerate}[resume]


\item Solve the Euler equation $x^2y'' - xy' - 3y = 0$ on the interval $(0;\infty)$ by using the substituion $x=e^t$ or otherwise. The answer should be written as a function of $x$.

\item Three players play the following game. Simulteneously each of them chooses one possible bet: either 1\$ or 2\$. A player is declared winner if his bet is unique and wins one rouble independently of the amount of the bet. For example, if players have chosen 1, 2 and 1 then their corresponding payoffs are 0, 1 and 0.

\begin{enumerate}
  \item Find all Nash equilibria in pure strategies.
  \item Find symmetric Nash equilibrium in mixed strategies.
\end{enumerate}

\end{enumerate}





\end{document}
